### 2018-01-31 includes/abstract.tex
Adding citation for FuzzSAT. Fixing up more content.

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem instance
    transformer and generator for string solvers. We supply a
    repository of instances generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    instances, and identify instances that are particularly challenging for
    some solvers. We briefly explain our observations and show how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}

>>>>

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem instance
    transformer and generator for string solvers. We supply a
    repository of instances generated by \fuzzer{} in \smtfull{}
    format. We systematically compare \theSolvers{} on groups of such
    instances, and identify the ones that are particularly challenging for
    some solvers. We briefly explain our observations and show how
    \fuzzer{} helped find causes of performance degradations in
    \us{}.
\end{abstract}

### 2018-01-31 includes/experiment.tex
Adding citation for FuzzSAT. Fixing up more content.

\fuzzer{}'s ability to produce scaling instances helped uncover several
implementation issues and performance limitations in \us{}. Scaling inputs
can reveal issues that would be out of scope for unit tests or
industrial benchmarks because they elicit asymptotic behaviour. Three different
performance and implementation bugs were identified and fixed in \us{}
as a result of testing with the \fuzzer{} suites
\textit{Lengths} and \textit{Concats}.

We also found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{Concats} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As an immediate consequence of
testing \us{} on \textit{Concats} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly.

>>>>

\fuzzer{}'s ability to produce scaling instances helped uncover several
implementation issues and performance limitations in \us{}. Scaling inputs
can reveal issues that would be out of scope for unit tests or
industrial benchmarks because they elicit asymptotic behaviour. Three different
performance and implementation bugs were identified and fixed in \us{}
as a result of testing with the \fuzzer{} scaling suites
\textit{Lengths} and \textit{Concats}.

We also found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \textit{Concats} suite helped us discover a
potential new heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations. As a consequence of
testing \us{} with the \textit{Concats} suite, we were able to
identify the factors common to its instances, examine \us{}'s execution
traces on them, and rapidly formulate a hypothesis as to why \us{}
was performing poorly.

### 2018-01-31 includes/introduction.tex
Adding citation for FuzzSAT. Fixing up more content.

In recent years many algorithms for solving string constraints have
been developed and implemented in SMT solvers such as Norn~\cite{norn},
CVC4~\cite{cvc4}, and Z3 (e.g., Z3str2~\cite{z3str2} and Z3str3~\cite{z3str3}).
To validate and benchmark these solvers, their developers have relied on
hand-crafted input suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or
real-world examples from a limited class of industrial
applications~\cite{kaluza,kausler}. These test suites have helped
developers identify implementation defects and develop more
sophisticated solving heuristics. Unfortunately, as solvers grow,
fixed benchmarks leave more functionality untested.
There is an acute need for a more robust, inexpensive, and automatic way
of generating benchmarks to test
the correctness and performance of SMT solvers.

Fuzzing is often used to test all kinds of software,
including SAT solvers. Inspired by their utility, we introduce a
fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate its efficacy
by presenting limitations it helped us find in
leading string solvers. To the best of our knowledge, \fuzzer{} is the
first tool aimed at automatic generation of string constraints. \fuzzer{} can
be used to mutate or transform existing benchmarks, as well as
randomly generate structured instances. These instances can be scaled with
respect to a variety of parameters, e.g., length of string constants,
depth of concats and regeular expressions, number of variables,
number of length constraints, and many more.

@@

\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In section~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and randomly generate \smtfull{} string and regular
        expression (regex) instances.\footnote{We assume basic
        familiarity with string solvers and their input
        language.} Scaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate such
        scaling inputs. We briefly document \fuzzer{}'s
        components and modular architecture. We provide example use cases to
        demonstrate its utility as an exploratory tool for validating
        solvers.

@@

    \item \textbf{Experimental Results and Analysis}:
        We compare the performance of \theSolvers{} on the
        the \fuzzer{} suites \theSuites{} in section~\ref{sec:data}. We
        highlight these suites because they make some solvers perform poorly,
        but not others. We analyze our
        experimental results, and pinpoint algorithmic limitations
        in \us{} that cause poor performance. We propose enhancements that
        address these issues, which \fuzzer{} helped us find.
\end{enumerate}

>>>>

In recent years, many algorithms for solving string constraints have
been developed and implemented in SMT solvers such as Norn~\cite{norn},
CVC4~\cite{cvc4}, and Z3 (e.g., Z3str2~\cite{z3str2} and Z3str3~\cite{z3str3}).
To validate and benchmark these solvers, their developers have relied on
hand-crafted input suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or
real-world examples from a limited set of industrial
applications~\cite{kaluza,kausler}. These test suites have helped
developers identify implementation defects and develop more
sophisticated solving heuristics. Unfortunately, as solvers grow,
fixed benchmarks leave more functionality untested.
There is an acute need for a more robust, inexpensive, and automatic way
of generating benchmarks to test
the correctness and performance of SMT solvers.

Fuzzing is often used to test all kinds of software,
including SAT solvers~\cite{fuzzsat}. Inspired by the utility of fuzzers,
we introduce a fuzzer for string solvers, \fuzzer{}, and describe its value
as an exploratory testing tool. We demonstrate its efficacy
by presenting limitations it helped us find in
leading string solvers. To the best of our knowledge, \fuzzer{} is the
first tool aimed at automatic generation of string constraints. \fuzzer{} can
be used to mutate or transform existing benchmarks, as well as
randomly generate structured instances. These instances can be scaled with
respect to a variety of parameters, e.g., length of string constants,
depth of concats and regeular expressions, number of variables,
number of length constraints, and many more.

@@

\begin{enumerate}
    \item \textbf{The \fuzzer{} tool}:
        In section~\ref{sec:fuzzer}, we describe a modular fuzzer that can
        transform and generate \smtfull{} string and regular
        expression (regex) instances.\footnote{We assume basic
        familiarity with string solvers and their input
        language.} Scaling inputs (e.g., long string constants,
        deep concatenations) are particularly useful in identifying asymptotic
        behaviors in solvers, and \fuzzer{} has many options to generate them.
        We briefly document \fuzzer{}'s components and modular architecture.
        We provide example use cases to demonstrate its utility as an
        exploratory solver testing tool.

@@

    \item \textbf{Experimental Results and Analysis}:
        We compare the performance of \theSolvers{} on the
        the \fuzzer{} suites \theSuites{} in section~\ref{sec:data}. We
        highlight these suites because they make some solvers perform poorly,
        but not others. We analyze our
        experimental results, and pinpoint algorithmic limitations
        in \us{} that cause poor performance.
\end{enumerate}

### 2018-01-31 includes/prior.tex
Adding citation for FuzzSAT. Fixing up more content.

Many solver developers author their own test suites to
validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. In
addition, several popular instance suites are publicly available for
solver testing and benchmarking, such as the Kaluza~\cite{kaluza} and
Kausler~\cite{kausler} suites. There are likewise several fuzzers and
instance generators currently available, but none of them can generate
or transform string and regex instances. For example, the
FuzzSMT~\cite{fuzzsmt} tool generates \smt{} instances with bit-vectors
and arrays, but does not support strings or regexes. The
SMTpp~\cite{smtpp} tool pre-processes and simplifies instances, but
does not generate new ones or fuzz existing ones.

>>>>

Many solver developers create their own test suites to
validate their solvers~\cite{cvc4-tests,z3str3-tests,z3str2-tests}. Several
popular instance suites are also publicly available for
solver testing and benchmarking, such as the Kaluza~\cite{kaluza} and
Kausler~\cite{kausler} suites. There are likewise several fuzzers and
instance generators currently available, but none of them can generate
or transform string and regex instances. For example, the
FuzzSMT~\cite{fuzzsmt} tool generates \smt{} instances with bit-vectors
and arrays, but does not support strings or regexes. The
SMTpp~\cite{smtpp} tool pre-processes and simplifies instances, but
does not generate new ones or fuzz existing ones.

### 2018-01-31 includes/stringfuzz.tex
Adding citation for FuzzSAT. Fixing up more content.

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. To demonstrate
its modularity, we note that the implementations of transformers in \fuzzer{}
contain an average of \linesPerX{} lines of code, whereas the whole fuzzer
contains \linesInFuzzer{} lines.

Its components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} instances. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        instances. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{Translate} and \textit{Reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        instance as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

@@

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
instance is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:
\begin{align*}
  & \texttt{(re.++}\; \ T_0\; \ T_1\; \texttt{...}\; T_n\; \texttt{)}
\end{align*}

@@

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are three example regexes
(separated by spaces) of depth 2 produced by this scheme:
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad\quad
  ((\texttt{ddd})*)+\quad\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

@@

\fuzzer{} can be used to generate benchmark suites in a batch mode, but also
in an online randomized fashion for exploratory debugging. For example,
\fuzzer{} can be invoked as follows to feed random instances to \cvc{} until it
returns an error:
{\scriptsize\begin{verbatim}
while stringfuzzg -r random-ast -m | tee instance.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done\end{verbatim}}

Simply by running the above script, we managed to induce segmentation faults
in \cvc{} and \us{}.

>>>>

\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string and regex
instances. It can either be installed from source, or from the Python
PIP package repository.\footnote{The link to the
source code will be added after double-blind review.}
\fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. To demonstrate
its modularity, we note that the implementations of transformers in \fuzzer{}
contain \linesPerX{} lines of code on average, whereas the whole fuzzer
contains \linesInFuzzer{} lines. Its components are also
implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver. \fuzzer{} is composed of the following tools:
\begin{description}
    \item[\generator{}] \hfill \\
        This tool generates \smt{} instances. It supports several generators and
        options that specify its output. Details can be found in
        table~\ref{tbl:generators}.
    \item[\transformer{}] \hfill \\
        This tool transforms \smt{}
        instances. It supports several transformers and options that specify
        its output and input, which are explained in
        table~\ref{tbl:transformers}. Note that transformers
        \textit{Translate} and \textit{Reverse} also preserve
        satisfiability under certain conditions~\cite{ifaz}.
    \item[\texttt{stringstats}] \hfill \\
        This tool takes an \smt{}
        instance as input and outputs its properties: the number of
        variables/literals, the max/median syntactic depth of expressions, the
        max/median literal length, etc.
\end{description}

@@

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
instance is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:
\begin{align*}
  & \texttt{(re.++}\; T_0\; \texttt{(re.++}\; T_1\;
  \texttt{...}\; \texttt{(re.++}\; T_{n-1}\; T_n\; \texttt{))}
\end{align*}

@@

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms, where each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are three example regexes
(separated by spaces) of depth 2 produced by this scheme:
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad\quad
  ((\texttt{ddd})*)+\quad\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}

@@

\fuzzer{} can be used to generate benchmark suites in a batch mode, but also
in an online randomized fashion for exploratory debugging. For example,
\fuzzer{} can be invoked as follows to feed random instances to \cvc{} until it
returns an error:
{\scriptsize\begin{verbatim}
while stringfuzzg -r random-ast -m \
    | tee instance.smt25 | cvc4 --lang smt2.5 --tlimit=5000 --strings-exp; do
    sleep 0
done\end{verbatim}}
